---
title: "Code Signing and Distribution"
vignette: >
  %\VignetteIndexEntry{Code Signing and Distribution}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

```{r}
#| include: false
library(shinyelectron)
```

Signing tells the OS who built this. Notarization tells the OS that Apple has seen it. Without either, macOS Gatekeeper blocks your app as an "unidentified developer" and Windows SmartScreen warns on every download. This guide covers what you need, what it costs, and how to wire it up. For the wider security picture, see [Security Considerations](security.html).

## Why sign at all

Users can click past warnings. The reasons to sign anyway:

- **Trust.** Non-technical users read "unidentified developer" as "malware" and leave.
- **Enterprise.** Many IT policies block anything unsigned.
- **Auto-updates.** `electron-updater` verifies signatures on macOS and Windows. Unsigned means no updates.

| | macOS | Windows | Linux |
|---|---|---|---|
| **Required?** | Strongly recommended | Strongly recommended | Optional |
| **Cost** | $99/year (Apple Developer) | $200 to $700/year (CA cert) | Free (GPG) |
| **Certificate** | Developer ID Application | OV or EV code signing | GPG key |
| **Notarization** | Yes (macOS 10.15+) | N/A | N/A |

## macOS: sign and notarize

### What you need

1. **Apple Developer Program membership** ($99/year): [developer.apple.com](https://developer.apple.com/programs/).
2. **Developer ID Application certificate**. Not the same as a Mac App Store certificate. Create it under Certificates, Identifiers & Profiles.
3. **App-specific password** for notarization, generated at [appleid.apple.com](https://appleid.apple.com/account/manage).

### Signing steps

Export the Developer ID Application certificate from Keychain Access as a `.p12` file, then set these variables:

```bash
# Certificate (path to .p12 file, or base64-encoded content)
export CSC_LINK="/path/to/certificate.p12"
export CSC_KEY_PASSWORD="your-certificate-password"

# Notarization credentials
export APPLE_ID="your@apple.id"
export APPLE_APP_SPECIFIC_PASSWORD="xxxx-xxxx-xxxx-xxxx"
export APPLE_TEAM_ID="XXXXXXXXXX"
```

Then export with signing on:

```{r}
#| eval: false
export(
  appdir = "my-app",
  destdir = "build",
  sign = TRUE
)
```

electron-builder drives the signing and notarization flow end to end. You do not need to call `codesign` or `xcrun notarytool` yourself.

You can also set signing options in `_shinyelectron.yml`. Keep credentials out of the config file: use environment variables.

```yaml
signing:
  sign: true
  mac:
    identity: "Developer ID Application: Your Name (TEAMID)"
    team_id: "XXXXXXXXXX"
    notarize: true
```

To open an unsigned `.app` during local development, strip the quarantine flag with `xattr -cr /path/to/YourApp.app` or right-click and choose **Open**.

## Windows: sign with a code signing certificate

### What you need

An OV or EV code signing certificate from a CA (DigiCert, Sectigo, GlobalSign, SSL.com). An EV certificate (~$300 to $700/year) gives you SmartScreen trust immediately. An OV certificate (~$200 to $500/year) builds reputation over weeks: users will still see warnings at first.

### Signing steps

```bash
# Path to your .pfx certificate file (or base64-encoded content)
export CSC_LINK="/path/to/certificate.pfx"
export CSC_KEY_PASSWORD="your-certificate-password"
```

Then build:

```{r}
#| eval: false
export(
  appdir = "my-app",
  destdir = "build",
  platform = "win",
  sign = TRUE
)
```

Unsigned Windows installers trigger SmartScreen on every download. There is no workaround short of signing.

## Linux: optional GPG signing

Linux distributions do not require code signing. AppImage files can still be GPG-signed so users who care can verify authenticity.

```bash
export GPG_KEY="your-gpg-key-id"
```

```yaml
signing:
  sign: true
  linux:
    gpg_sign: true
```

Most users never check. Sign if your audience does.

## The `sign` parameter

`export(..., sign = TRUE)` tells electron-builder to use whatever credentials it finds. It writes signing configuration into `package.json`, sets the macOS identity, enables notarization (if configured), and points Windows at the certificate. It does **not** create certificates. If credentials are missing, it warns and hands off to electron-builder, which may then fail at the signing step.

The function argument wins over the config file. `export(..., sign = TRUE)` signs even if `_shinyelectron.yml` has `sign: false`. You can also enable signing purely through the config (`signing: sign: true`).

Run `app_check("my-app", sign = TRUE)` to verify your setup before you kick off a full build.

## GitHub Actions

Store credentials as GitHub Actions secrets (**Settings > Secrets and variables > Actions**) and reference them from your workflow.

| Secret Name | Value |
|------------|-------|
| `MAC_CERTIFICATE` | Base64-encoded `.p12` file (`base64 -i cert.p12`) |
| `MAC_CERTIFICATE_PASSWORD` | Password for the `.p12` file |
| `APPLE_ID` | Your Apple ID email |
| `APPLE_APP_SPECIFIC_PASSWORD` | App-specific password from appleid.apple.com |
| `APPLE_TEAM_ID` | 10-character team ID from developer.apple.com |
| `WIN_CERTIFICATE` | Base64-encoded `.pfx` file |
| `WIN_CERTIFICATE_PASSWORD` | Password for the `.pfx` file |

Then in the workflow:

```yaml
- name: Build Electron app
  env:
    # macOS
    CSC_LINK: ${{ secrets.MAC_CERTIFICATE }}
    CSC_KEY_PASSWORD: ${{ secrets.MAC_CERTIFICATE_PASSWORD }}
    APPLE_ID: ${{ secrets.APPLE_ID }}
    APPLE_APP_SPECIFIC_PASSWORD: ${{ secrets.APPLE_APP_SPECIFIC_PASSWORD }}
    APPLE_TEAM_ID: ${{ secrets.APPLE_TEAM_ID }}
    # Windows
    # CSC_LINK: ${{ secrets.WIN_CERTIFICATE }}
    # CSC_KEY_PASSWORD: ${{ secrets.WIN_CERTIFICATE_PASSWORD }}
  run: |
    Rscript -e "
      shinyelectron::export(
        appdir = 'app',
        destdir = 'build',
        sign = TRUE
      )
    "
```

::: {.callout-important}
macOS and Windows use different certificates. `CSC_LINK` and `CSC_KEY_PASSWORD` must point at the right one for the runner. In a matrix build, set them per-platform with conditional env blocks or split the build steps.
:::

For a full workflow template, see [GitHub Actions](github-actions.html).

## Development versus production

| | Development | Production |
|---|---|---|
| **`sign`** | `FALSE` (default) | `TRUE` |
| **Credentials needed?** | No | Yes |
| **Build speed** | Fast | Slower: notarization takes 1 to 5 minutes |
| **OS warnings?** | Yes (use `xattr -cr` on macOS) | No |

A working rhythm:

1. Develop and iterate with `sign = FALSE`.
2. Set up signing credentials once.
3. Flip `sign = TRUE` (or `signing: sign: true`) for release builds.
4. Push signing to CI so humans never hold the keys locally.

## Quick reference

| Environment Variable | Platform | Purpose |
|---------------------|----------|---------|
| `CSC_LINK` | macOS, Windows | Path to `.p12`/`.pfx` certificate (or base64) |
| `CSC_KEY_PASSWORD` | macOS, Windows | Certificate password |
| `APPLE_ID` | macOS | Apple ID for notarization |
| `APPLE_APP_SPECIFIC_PASSWORD` | macOS | App-specific password for notarization |
| `APPLE_TEAM_ID` | macOS | 10-character Apple Developer Team ID |
| `GPG_KEY` | Linux | GPG key ID for AppImage signing |

## Next steps

- **[GitHub Actions](github-actions.html)**: automate signed builds in CI.
- **[Auto Updates](auto-updates.html)**: auto-updates require signed builds on macOS and Windows.
- **[Configuration](configuration.html)**: the full `_shinyelectron.yml` reference, signing included.
