---
title: "Configuration Guide"
vignette: >
  %\VignetteIndexEntry{Configuration Guide}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

```{r}
#| include: false
library(shinyelectron)
```

The shinyelectron package supports a YAML configuration file that simplifies
the export process by storing settings in your project directory. This guide
explains the configuration file format and all available options.

## Why Use a Configuration File?

Instead of passing many parameters to `export()`:

```{r}
#| eval: false
export(
  appdir = "my-app",
  destdir = "output",
  app_name = "My Application",
  app_type = "r-shinylive",
  platform = c("mac", "win"),
  arch = c("x64", "arm64"),
  icon = "icons/icon.icns"
)
```
You can store these settings in a `_shinyelectron.yml` file and simply run:

```{r}
#| eval: false
export(appdir = "my-app", destdir = "output")
```

The configuration file also:

- Keeps build settings version-controlled with your project
- Allows sharing settings across team members
- Documents your project's build configuration

## Creating a Configuration File

Use `init_config()` to create a template configuration file:

```{r}
#| eval: false
init_config("path/to/my-app")
```
```
✔ Created configuration file: path/to/my-app/_shinyelectron.yml
ℹ Edit this file to customize your Electron app settings
```

This creates a `_shinyelectron.yml` file with documented defaults.

## Configuration File Location

The configuration file must be named `_shinyelectron.yml` and placed in the
root of your Shiny application directory:

```
my-shiny-app/
├── _shinyelectron.yml    # Configuration file
├── app.R                 # Your Shiny app
└── ...
```

## Complete Configuration Reference

Here is a complete configuration file with all available options:

```yaml
# shinyelectron configuration file
# Documentation: https://r-pkg.thecoatlessprofessor.com/shinyelectron/

app:
  name: "My Application"     # Application display name
  version: "1.0.0"           # Application version (used in package metadata)

build:
  type: "r-shinylive"        # Application type
  runtime_strategy: "auto-download"  # For native types: bundled, system, auto-download, container
  platforms:                 # Target platforms
    - mac
    - win
    - linux
  architectures:             # Target architectures
    - x64
    - arm64

r:
  version: null              # R version for bundled/auto-download (null = latest release)

python:
  version: null              # Python version for bundled/auto-download (null = "3.12.10")

window:
  width: 1200                # Default window width (pixels)
  height: 800                # Default window height (pixels)

server:
  port: 3838                 # Development server port

icons:                       # Platform-specific icons
  mac: "icons/icon.icns"
  win: "icons/icon.ico"
  linux: "icons/icon.png"

nodejs:
  version: null              # Node.js version (null = latest LTS)
  auto_install: false        # Auto-install Node.js if not found

container:                   # Used when runtime_strategy is "container"
  engine: "docker"           # "docker" or "podman"
  image: null                # Container image (null = auto-select)
  tag: "latest"
  pull_on_start: true        # Pull latest image on app start
  volumes: []                # Additional volume mounts
  env: []                    # Additional environment variables

# Multi-app suite (2+ apps packaged in one Electron shell)
# apps:
#   - id: "dashboard"
#     name: "Dashboard"
#     path: "./apps/dashboard"
#     type: "r-shiny"          # Optional per-app override (default: build.type)
#     description: "Main dashboard"
#     icon: "icons/dash.png"
#   - id: "admin"
#     name: "Admin Panel"
#     path: "./apps/admin"
```

## Configuration Sections

### `app` Section

Controls application metadata.

| Key | Type | Default | Description |
|-----|------|---------|-------------|
| `name` | string | Directory name | Display name shown in window title and system |
| `version` | string | `"1.0.0"` | Version number for the built application |

**Example:**

```yaml
app:
  name: "Sales Dashboard"
  version: "2.1.0"
```

### `build` Section

Controls the build process and targets.

| Key | Type | Default | Description |
|-----|------|---------|-------------|
| `type` | string | `"r-shinylive"` | Application type (see below) |
| `runtime_strategy` | string | `"auto-download"` | Runtime strategy for native types (see below) |
| `platforms` | list | Current platform | Target operating systems |
| `architectures` | list | Current arch | Target CPU architectures |

**Valid `type` values:**

| Type | Description |
|------|-------------|
| `r-shinylive` | R Shiny app converted to run in browser via WebR |
| `py-shinylive` | Python Shiny app converted to run in browser via Pyodide |
| `r-shiny` | Native R Shiny app with runtime |
| `py-shiny` | Native Python Shiny app with runtime |

**Valid `runtime_strategy` values** (native types only):

| Strategy | Description |
|----------|-------------|
| `auto-download` | Downloads R/Python on first launch, caches for reuse (default) |
| `bundled` | Embeds a portable R/Python runtime in the app at build time |
| `system` | Uses R/Python already installed on the end user's machine |
| `container` | Runs the app inside a Docker/Podman container |

**Valid `platforms` values:** `mac`, `win`, `linux`

**Valid `architectures` values:** `x64`, `arm64`

**Example:**

```yaml
build:
  type: "r-shiny"
  runtime_strategy: "bundled"
  platforms:
    - mac
    - win
  architectures:
    - x64
    - arm64
```

::: {.callout-note}
## Cross-Platform Building

Building for a platform different from your development machine may have
limitations. macOS apps can only be built on macOS. Windows and Linux apps can
be cross-compiled with additional setup.
:::

### `window` Section

Controls the Electron window dimensions.

| Key | Type | Default | Description |
|-----|------|---------|-------------|
| `width` | integer | `1200` | Window width in pixels (minimum: 100) |
| `height` | integer | `800` | Window height in pixels (minimum: 100) |

**Example:**

```yaml
window:
  width: 1400
  height: 900
```

### `server` Section

Controls development server settings.

| Key | Type | Default | Description |
|-----|------|---------|-------------|
| `port` | integer | `3838` | Port for local development server (1-65535) |

**Example:**

```yaml
server:
  port: 8080
```

### `icons` Section

Specifies custom application icons for each platform. Icons must be in
platform-specific formats.

| Key | Type | Format | Description |
|-----|------|--------|-------------|
| `mac` | string | `.icns` | macOS icon (typically 512x512 or larger) |
| `win` | string | `.ico` | Windows icon (multi-resolution recommended) |
| `linux` | string | `.png` | Linux icon (512x512 recommended) |

Paths are relative to your app directory.

**Example:**

```yaml
icons:
  mac: "assets/app-icon.icns"
  win: "assets/app-icon.ico"
  linux: "assets/app-icon.png"
```

::: {.callout-tip}
## Creating Icons

Tools like [iconutil](https://developer.apple.com/library/archive/documentation/GraphicsAnimation/Conceptual/HighResolutionOSX/Optimizing/Optimizing.html)
(macOS), [ImageMagick](https://imagemagick.org/), or online converters can
create platform-specific icon formats from a single high-resolution PNG.
:::

### `nodejs` Section

Controls Node.js installation behavior.

| Key | Type | Default | Description |
|-----|------|---------|-------------|
| `version` | string | `null` | Node.js version to install (`null` = latest LTS) |
| `auto_install` | boolean | `false` | Auto-install Node.js if not found |

**Example:**

```yaml
nodejs:
  version: "22.0.0"
  auto_install: true
```

::: {.callout-warning}
## Auto-Install Opt-In

Auto-install is disabled by default. Set `auto_install: true` explicitly to
enable automatic Node.js installation during `export()`.
:::

### `r` Section

Pins the R version used for `bundled` and `auto-download` runtime strategies.

| Key | Type | Default | Description |
|-----|------|---------|-------------|
| `version` | string | `null` | R version (e.g., `"4.4.1"`). `null` = latest release. |

**Example:**

```yaml
r:
  version: "4.4.1"
```

### `python` Section

Pins the Python version used for `bundled` and `auto-download` runtime strategies.

| Key | Type | Default | Description |
|-----|------|---------|-------------|
| `version` | string | `null` | Python version (e.g., `"3.12.10"`). `null` = `"3.12.10"`. |

**Example:**

```yaml
python:
  version: "3.12.10"
```

### `container` Section

Configures the container backend when `runtime_strategy` is `"container"`.

| Key | Type | Default | Description |
|-----|------|---------|-------------|
| `engine` | string | `"docker"` | Container engine: `"docker"` or `"podman"` |
| `image` | string | `null` | Container image name (`null` = auto-select) |
| `tag` | string | `"latest"` | Image tag |
| `pull_on_start` | boolean | `true` | Pull the latest image when the app starts |
| `volumes` | list | `[]` | Additional volume mounts |
| `env` | list | `[]` | Additional environment variables |

**Example:**

```yaml
container:
  engine: "docker"
  image: "rocker/shiny"
  tag: "4.4.1"
  pull_on_start: true
  volumes:
    - "/data:/app/data:ro"
  env:
    - "SHINY_LOG_LEVEL=DEBUG"
```

### `apps` Section

Defines a multi-app suite -- two or more Shiny apps packaged in a single
Electron shell with a launcher screen. Requires at least two entries. Each app
can override the default `build.type`.

| Key | Type | Required | Description |
|-----|------|----------|-------------|
| `id` | string | yes | Unique identifier (URL-safe) |
| `name` | string | yes | Display name in launcher |
| `path` | string | yes | Relative path to app directory |
| `type` | string | no | App type override (default: `build.type`) |
| `description` | string | no | Short description shown in launcher |
| `icon` | string | no | Per-app icon path |

**Example:**

```yaml
apps:
  - id: "dashboard"
    name: "Dashboard"
    path: "./apps/dashboard"
    type: "r-shiny"
    description: "Sales analytics dashboard"
  - id: "admin"
    name: "Admin Panel"
    path: "./apps/admin"
    type: "py-shiny"
    description: "User management"
```

## Configuration Precedence

When you call `export()` with both a config file and function parameters,
values are merged with this precedence (highest to lowest):
1. **Function parameters** - Values passed directly to `export()`
2. **Configuration file** - Values from `_shinyelectron.yml`
3. **Default values** - Built-in defaults

**Example:**

```yaml
# _shinyelectron.yml
app:
  name: "Config Name"
build:
  platforms:
    - mac
```

```{r}
#| eval: false
# This uses "Override Name" (function param wins)
# and builds for mac only (from config)
export(
  appdir = "my-app",
  destdir = "output",
  app_name = "Override Name"
)
```

## Minimal Configuration

For most projects, a minimal configuration is sufficient:

```yaml
app:
  name: "My App"

nodejs:
  auto_install: true
```

All other values use sensible defaults.

## Common Configurations

### Development Setup

Quick iteration during development:

```yaml
app:
  name: "My App (Dev)"
  version: "0.0.1"

window:
  width: 1000
  height: 700

nodejs:
  auto_install: true
```

### Production Multi-Platform Build

Build for distribution across platforms:

```yaml
app:
  name: "Production App"
  version: "1.0.0"

build:
  type: "r-shinylive"
  platforms:
    - mac
    - win
    - linux
  architectures:
    - x64
    - arm64

window:
  width: 1200
  height: 800

icons:
  mac: "branding/icon.icns"
  win: "branding/icon.ico"
  linux: "branding/icon.png"

nodejs:
  version: "22.11.0"
  auto_install: false
```

### Native R App with Bundled Runtime

Package an R Shiny app with the runtime embedded:

```yaml
app:
  name: "Analytics Tool"
  version: "1.0.0"

build:
  type: "r-shiny"
  runtime_strategy: "bundled"
  platforms:
    - mac
    - win

r:
  version: "4.4.1"

nodejs:
  auto_install: true
```

### Multi-App Configuration

Package multiple Shiny apps in a single Electron shell with a launcher:

```yaml
app:
  name: "My App Suite"
  version: "1.0.0"

build:
  type: "r-shiny"
  runtime_strategy: "auto-download"

apps:
  - id: "dashboard"
    name: "Dashboard"
    path: "./apps/dashboard"
    description: "Sales analytics"
  - id: "admin"
    name: "Admin Panel"
    path: "./apps/admin"
    type: "py-shiny"
    description: "User management"

nodejs:
  auto_install: true
```

### Locked Node.js Version

Pin Node.js version for reproducible builds:

```yaml
nodejs:
  version: "22.11.0"
  auto_install: true
```

## Validation and Error Handling

Configuration values are validated when read:

- Invalid `type` values warn and fall back to `"r-shinylive"`
- Invalid platforms/architectures are filtered out with a warning
- Window dimensions below 100 pixels warn and use defaults
- Invalid port numbers warn and use 3838

If the YAML file has syntax errors, shinyelectron warns and uses all defaults.

## Next Steps

- **[Getting Started](getting-started.html)**: First-time user walkthrough
- **[Node.js Management](nodejs-management.html)**: Managing local Node.js
- **[Troubleshooting](troubleshooting.html)**: Diagnosing issues
