---
title: "Download Prebuilt Demos"
vignette: >
  %\VignetteIndexEntry{Download Prebuilt Demos}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

```{r}
#| include: false
library(shinyelectron)
```

Every demo is built with every runtime strategy on every supported platform and published as a release asset. Pick a row, check what it needs, and download.

What each strategy needs on the user's machine:

- `shinylive`: nothing. The app runs in WebAssembly.
- `bundled`: nothing. A runtime is embedded in the installer.
- `system`: R or Python already installed and on the PATH.
- `auto-download`: internet access on first launch to fetch the runtime.
- `container`: Docker or Podman installed.

Your platform's downloads are shown first. Use **More downloads** for every other platform.

```{r}
#| echo: false
#| results: asis
m <- shinyelectron:::demo_release_matrix()
base <- "https://github.com/coatless-rpkg/shinyelectron/releases/latest/download/"

platforms <- list(
  list(id = "mac-arm64", label = "macOS (Apple Silicon)", platform = "mac", arch = "arm64"),
  list(id = "mac-x64",   label = "macOS (Intel)",         platform = "mac", arch = "x64"),
  list(id = "win-x64",   label = "Windows",               platform = "win", arch = "x64"),
  list(id = "linux-x64", label = "Linux",                 platform = "linux", arch = "x64")
)
languages <- list(list(name = "Python", lang = "py"), list(name = "R", lang = "r"))
groups <- list(
  list(title = "Standalone (no setup)", strategies = c("shinylive", "bundled")),
  list(title = "Needs a runtime or tooling", strategies = c("system", "auto-download", "container"))
)

dl <- function(rows, demo, strat) {
  r <- rows[rows$demo == demo & rows$strategy == strat, ]
  if (nrow(r) == 0) return("--")
  sprintf("[download](%s%s)", base, r$asset_name)
}

for (p in platforms) {
  cat(sprintf("\n::: {.platform-block data-platform=\"%s\" data-label=\"%s\"}\n", p$id, p$label))
  pm <- m[m$platform == p$platform & m$arch == p$arch, ]
  for (lg in languages) {
    cat(sprintf("\n## %s\n", lg$name))
    lm <- pm[pm$language == lg$lang, ]
    demos <- unique(lm[, c("demo", "name")])
    for (g in groups) {
      strs <- g$strategies
      cat(sprintf("\n### %s\n\n", g$title))
      cat(paste0("| Demo | ", paste(sprintf("`%s`", strs), collapse = " | "), " |\n"))
      cat(paste0("|", paste(rep("---", length(strs) + 1), collapse = "|"), "|\n"))
      for (i in seq_len(nrow(demos))) {
        cells <- vapply(strs, function(s) dl(lm, demos$demo[i], s), character(1))
        cat(sprintf("| %s | %s |\n", demos$name[i], paste(cells, collapse = " | ")))
      }
    }
  }
  cat("\n:::\n")
}
```
