---
title: "Getting Started with shinyelectron"
vignette: >
  %\VignetteIndexEntry{Getting Started with shinyelectron}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

```{r}
#| include: false
library(shinyelectron)
```

The shinyelectron package converts Shiny applications (R and Python) into
standalone desktop applications using Electron. This guide walks you through
the complete workflow from installation to running your first desktop Shiny app.

## Prerequisites

Before using shinyelectron, ensure you have:

- **R** (>= 4.4.0)
- **Node.js** (>= 22.0.0) - Can be installed automatically via shinyelectron
- **npm** (>= 11.5.0) - Included with Node.js

For building platform-specific installers, you also need:

| Platform | Requirements |
|----------|-------------|
| Windows | Windows 11+, Visual Studio Build Tools |
| macOS | macOS 10.13+, Xcode Command Line Tools |
| Linux | gcc, make (build-essential package) |

## Installation

Install shinyelectron from GitHub:

```{r}
#| eval: false
# install.packages("remotes")
remotes::install_github("coatless-rpkg/shinyelectron")
```

## Step 1: Check System Readiness

Before exporting your first app, run the system diagnostics to verify everything
is properly configured:

```{r}
#| eval: false
library(shinyelectron)

# Run full system diagnostics
sitrep_shinyelectron()
```
```
── shinyelectron System Report ───────────────────────────────────────────────
✔ Platform: darwin (arm64)
✔ Node.js: v22.11.0 (local installation)
✔ npm: 10.9.0
✔ R packages: All required packages installed
ℹ Build tools: Xcode CLI available
```

This function checks:

- Node.js and npm availability
- R package dependencies
- Platform-specific build tools
- Local Node.js installation status

## Step 2: Install Node.js (Optional)

If Node.js is not installed on your system, shinyelectron can install it locally
without requiring administrator privileges:
```{r}
#| eval: false
# Install latest LTS version of Node.js
install_nodejs()
```
```
ℹ Detecting latest Node.js LTS version...
✔ Latest LTS: v22.11.0
── Installing Node.js v22.11.0 ───────────────────────────────────────────────
ℹ Platform: mac
ℹ Architecture: arm64
ℹ Fetching checksums...
ℹ Downloading from https://nodejs.org/dist/v22.11.0/node-v22.11.0-darwin-arm64.tar.gz
✔ Checksum verified
ℹ Extracting archive...
✔ Node.js v22.11.0 installed successfully
ℹ Location: ~/.shinyelectron/cache/assets/nodejs/v22.11.0/darwin-arm64
```

This installation is managed entirely by shinyelectron and does not affect any
system-wide Node.js installation you may have.

## Step 3: Create or Obtain a Shiny App

shinyelectron supports both R and Python Shiny apps. For this tutorial, we
start with the classic "Hello Shiny" example from the R shiny package:

```{r}
#| eval: false
# Copy the Hello Shiny example to a local directory
system.file("examples", "01_hello", package = "shiny") |>
  fs::dir_copy("my-shiny-app", overwrite = TRUE)
```

This creates a directory called `my-shiny-app` containing:

```
my-shiny-app/
├── app.R
└── (other files)
```

::: {.callout-tip}
## Using Your Own App

You can use any Shiny application directory. The expected entry point depends on
the language:

- **R Shiny**: `app.R` (single-file app), or both `server.R` and `ui.R`
- **Python Shiny**: `app.py`
:::

## Step 4: Export to Electron

The `export()` function handles the complete conversion process. The default
`app_type` is `"r-shinylive"`, which converts R Shiny apps to run entirely in
the browser via WebR -- no R installation required on the end user's machine.

### R Shinylive Export

```{r}
#| eval: false
export(
  appdir = "my-shiny-app",
  destdir = "my-electron-app",
  app_name = "Hello Shiny",
  run_after = TRUE,
  verbose = TRUE
)
```
```
── Exporting Shiny application to Electron ───────────────────────────────────
ℹ Application: Hello Shiny
ℹ Type: r-shinylive
ℹ Source: my-shiny-app
ℹ Destination: my-electron-app
ℹ Converting to shinylive format...
✔ Shinylive conversion complete
ℹ Building Electron application...
ℹ Setting up Electron project structure...
✔ Created project structure
ℹ Copying application files...
✔ Copied application files
ℹ Processing Electron templates...
✔ Processed Electron templates
ℹ Installing npm dependencies...
✔ Installed npm dependencies
ℹ Building for platforms: mac
✔ Built for mac-arm64
✔ Successfully exported Shiny application to Electron!
ℹ Output directory: my-electron-app
ℹ Electron app: my-electron-app/electron-app
ℹ Distributables: my-electron-app/electron-app/dist
```

The `run_after = TRUE` option automatically launches your app for testing after
the build completes.

### Python Shinylive Export

If you have a Python Shiny app (a directory containing `app.py`), set
`app_type = "py-shinylive"` to convert it via Pyodide:

```{r}
#| eval: false
export(
  appdir = "my-py-app",
  destdir = "my-py-electron-app",
  app_name = "Hello PyShiny",
  app_type = "py-shinylive"
)
```

The workflow is identical to the R shinylive case -- the app runs entirely in
the browser with no Python installation needed on the end user's machine.

## Native App Types

Shinylive exports are the simplest path, but they run inside a WebAssembly
sandbox and cannot use every R or Python package. For apps that need full
native R or Python, use `app_type = "r-shiny"` or `app_type = "py-shiny"`:

```{r}
#| eval: false
# Native R Shiny app (auto-downloads R on first launch)
export(
  appdir = "my-shiny-app",
  destdir = "my-native-electron-app",
  app_type = "r-shiny"
)

# Native Python Shiny app
export(
  appdir = "my-py-app",
  destdir = "my-native-py-electron-app",
  app_type = "py-shiny"
)
```

Native app types require a runtime on the end user's machine. The
`runtime_strategy` parameter controls how that runtime is provided:

| Strategy | Description |
|----------|-------------|
| `"auto-download"` | **(default)** Downloads R/Python on first launch; caches for reuse |
| `"bundled"` | Embeds a portable R/Python runtime inside the Electron app at build time |
| `"system"` | Uses R/Python already installed on the end user's machine |
| `"container"` | Runs the app inside a Docker or Podman container |

```{r}
#| eval: false
# Example: bundle R into the app so users need nothing pre-installed
export(
  appdir = "my-shiny-app",
  destdir = "my-bundled-app",
  app_type = "r-shiny",
  runtime_strategy = "bundled"
)
```

For a deeper look at trade-offs, setup, and configuration for each strategy,
see the [Runtime Strategies](runtime-strategies.html) article.

## Understanding the Output

After export, your destination directory contains:

```
my-electron-app/
├── shinylive-app/     # Converted shinylive application
└── electron-app/      # Complete Electron project
    ├── src/
    │   └── app/       # Your application files
    ├── main.js        # Electron main process
    ├── package.json   # npm configuration
    ├── node_modules/  # npm dependencies
    └── dist/          # Built distributables
        └── mac-arm64/
            └── Hello Shiny.app  # macOS application
```

### Finding Your Application

The final distributable application is in the `dist/` folder:

| Platform | Location | File Type |
|----------|----------|-----------|
| macOS | `dist/mac-arm64/` or `dist/mac-x64/` | `.app` bundle |
| Windows | `dist/win-x64/` | `.exe` installer |
| Linux | `dist/linux-x64/` | AppImage |

## Customizing Your Export

### Building for Multiple Platforms

Build for Windows, macOS, and Linux in a single command:

```{r}
#| eval: false
export(
 appdir = "my-shiny-app",
 destdir = "my-electron-app",
 platform = c("mac", "win", "linux"),
 arch = c("x64", "arm64"),
 overwrite = TRUE
)
```

::: {.callout-note}
Cross-platform building has limitations. Building Windows apps on macOS or Linux
requires additional tools, and building macOS apps requires macOS.
:::

### Adding a Custom Icon

Provide platform-specific icons:

```{r}
#| eval: false
export(
  appdir = "my-shiny-app",
  destdir = "my-electron-app",
  icon = "path/to/icon.icns"  # .icns for macOS, .ico for Windows
)
```

### Using a Configuration File

For complex projects, use a `_shinyelectron.yml` configuration file instead of
passing many parameters. See the [Configuration](configuration.html) article for
details.

## Next Steps

- **[Configuration](configuration.html)**: Learn about the `_shinyelectron.yml`
  configuration file for advanced customization
- **[Runtime Strategies](runtime-strategies.html)**: Deep dive into bundled,
  system, auto-download, and container strategies for native app types
- **[Multi-App Suites](multi-app-suites.html)**: Package multiple Shiny apps
  into a single Electron application
- **[Node.js Management](nodejs-management.html)**: Manage local Node.js
  installations
- **[Troubleshooting](troubleshooting.html)**: Diagnose and fix common issues

## Quick Reference

| Function | Purpose |
|----------|---------|
| `export()` | Convert and build Shiny app to Electron |
| `app_check()` | Pre-flight validation without building |
| `wizard()` | Interactive configuration generator |
| `show_config()` | Display merged effective configuration |
| `sitrep_shinyelectron()` | Run full system diagnostics |
| `install_nodejs()` | Install Node.js locally |
| `init_config()` | Create configuration file |
| `enable_auto_updates()` | Configure auto-update via GitHub Releases |
| `run_electron_app()` | Test an Electron app |
| `cache_clear()` | Clear cached assets |
