---
title: "Building with GitHub Actions"
vignette: >
  %\VignetteIndexEntry{Building with GitHub Actions}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

```{r}
#| include: false
library(shinyelectron)
```
GitHub Actions provides a powerful way to automatically build your Shiny
application as an Electron desktop app for Windows, macOS, and Linux. This
guide explains how to set up CI/CD for your shinyelectron project.

## Why Use GitHub Actions?

Building desktop applications manually for multiple platforms is time-consuming:

| Challenge | GitHub Actions Solution |
|-----------|------------------------|
| Need access to Windows, macOS, and Linux | Runners available for all platforms |
| Manual builds are error-prone | Automated, reproducible builds |
| Distributing builds is tedious | Automatic artifact upload and releases |
| Testing across platforms | Matrix builds run in parallel |

## Prerequisites

Before setting up GitHub Actions, ensure you have:

1. A GitHub repository containing your Shiny app
2. Your Shiny app in a subdirectory (default: `app/`)
3. (Optional) A `_shinyelectron.yml` configuration file

**Example repository structure:**

```
my-shiny-project/
├── .github/
│   └── workflows/
│       └── build-electron.yml   # Workflow file
├── app/
│   ├── app.R                    # Your Shiny app
│   └── ...
├── _shinyelectron.yml           # Optional config
└── README.md
```

## Quick Start

### Step 1: Copy the Template

shinyelectron includes a ready-to-use workflow template. Copy it to your
repository:

```{r}
#| eval: false
# Find the template location
template_path <- system.file(
 "templates",
 "github-actions-build.yml",
 package = "shinyelectron"
)

# View the template
file.show(template_path)

# Copy to your project (run from your project root)
dir.create(".github/workflows", recursive = TRUE, showWarnings = FALSE)
file.copy(
 template_path,
 ".github/workflows/build-electron.yml"
)
```

Or download directly from:
<https://github.com/coatless-rpkg/shinyelectron/blob/main/inst/templates/github-actions-build.yml>

### Step 2: Configure the Workflow

Edit `.github/workflows/build-electron.yml` and update the environment
variables at the top:

```yaml
env:
 APP_DIR: 'app'           # Directory containing your Shiny app
 APP_NAME: 'MyApp'        # Name of your application
 NODE_VERSION: '22'       # Node.js version
 R_VERSION: 'release'     # R version
```

### Step 3: Push and Build

Commit and push your changes:

```bash
git add .github/workflows/build-electron.yml
git commit -m "Add Electron build workflow"
git push
```
The workflow triggers automatically on push to main/master branches.

## Workflow Structure

The template workflow has three jobs:

### Build Job

Runs on each platform in parallel:

| Runner | Platform | Architecture | Output |
|--------|----------|--------------|--------|
| `macos-latest` | macOS | arm64 (Apple Silicon) | `.dmg` |
| `macos-13` | macOS | x64 (Intel) | `.dmg` |
| `windows-latest` | Windows | x64 | `.exe` |
| `ubuntu-latest` | Linux | x64 | `.AppImage` |

**Key steps:**

1. **Checkout** - Clone your repository
2. **Setup R** - Install R using r-lib/actions
3. **Setup Node.js** - Install Node.js 22
4. **Install dependencies** - R packages including shinyelectron
5. **Build** - Run `shinyelectron::export()`
6. **Upload artifacts** - Save build outputs

### Release Job

Triggers only when you push a version tag (e.g., `v1.0.0`):

- Downloads all build artifacts
- Creates a GitHub Release
- Attaches all platform builds

### Summary Job

Provides a build status summary in the Actions UI.

## Customization

### Changing the App Directory

If your Shiny app isn't in `app/`, update `APP_DIR`:

```yaml
env:
 APP_DIR: 'src/shiny-app'  # Custom location
```

### Building for Specific Platforms

To build only for certain platforms, modify the matrix:

```yaml
matrix:
 include:
   # Only macOS and Windows
   - os: macos-latest
     platform: mac
     arch: arm64
   - os: windows-latest
     platform: win
     arch: x64
```

### Using a Configuration File

If you have a `_shinyelectron.yml` in your app directory, the build will
automatically use it. The workflow parameters override config file values.

Example `app/_shinyelectron.yml`:

```yaml
app:
 name: "My Shiny Dashboard"
 version: "1.0.0"

window:
 width: 1400
 height: 900

build:
 type: "r-shinylive"
```

### Adding Custom Icons

Place icons in your repository and reference them:

```yaml
- name: Build Electron app
 run: |
   Rscript -e "
     shinyelectron::export(
       appdir = '${{ env.APP_DIR }}',
       destdir = 'build',
       icon = 'assets/icon.icns',
       # ... other options
     )
   "
```

::: {.callout-note}
Icon format requirements:

- macOS: `.icns` (use `iconutil` to create)
- Windows: `.ico` (multi-resolution recommended)
- Linux: `.png` (512x512 recommended)
:::

## Creating Releases

### Automatic Releases

Push a version tag to create a release automatically:

```bash
git tag v1.0.0
git push origin v1.0.0
```

The workflow will:

1. Build for all platforms
2. Create a GitHub Release named "v1.0.0"
3. Attach all build artifacts

### Pre-releases

Tags containing `-alpha` or `-beta` are marked as pre-releases:

```bash
git tag v1.0.0-beta.1
git push origin v1.0.0-beta.1
```

### Manual Releases

Trigger a build manually from the Actions tab using "workflow_dispatch":

1. Go to Actions → Build Electron App
2. Click "Run workflow"
3. Select branch and click "Run workflow"

## Advanced Topics

### Caching Dependencies

The template includes npm caching. To also cache R packages:

```yaml
- name: Setup R
 uses: r-lib/actions/setup-r@v2
 with:
   r-version: ${{ env.R_VERSION }}
   use-public-rspm: true

- name: Cache R packages
 uses: actions/cache@v4
 with:
   path: ${{ env.R_LIBS_USER }}
   key: ${{ runner.os }}-r-${{ hashFiles('**/DESCRIPTION') }}
```

### Build Status Badge

Add a badge to your README:

```markdown
[![Build Electron App](https://github.com/YOUR-USERNAME/YOUR-REPO/actions/workflows/build-electron.yml/badge.svg)](https://github.com/YOUR-USERNAME/YOUR-REPO/actions/workflows/build-electron.yml)
```

### Code Signing

Code signing requires additional setup:

**macOS:**

```yaml
env:
 CSC_LINK: ${{ secrets.MAC_CERTIFICATE }}
 CSC_KEY_PASSWORD: ${{ secrets.MAC_CERTIFICATE_PASSWORD }}
 APPLE_ID: ${{ secrets.APPLE_ID }}
 APPLE_APP_SPECIFIC_PASSWORD: ${{ secrets.APPLE_APP_SPECIFIC_PASSWORD }}
```

**Windows:**

```yaml
env:
 CSC_LINK: ${{ secrets.WIN_CERTIFICATE }}
 CSC_KEY_PASSWORD: ${{ secrets.WIN_CERTIFICATE_PASSWORD }}
```

::: {.callout-warning}
Code signing requires purchasing certificates from Apple and/or a Certificate
Authority. Unsigned apps may trigger security warnings on user machines.
:::

### Secrets Management

Store sensitive values in GitHub Secrets:

1. Go to Settings → Secrets and variables → Actions
2. Click "New repository secret"
3. Add your secret (e.g., `MAC_CERTIFICATE`)

Reference in workflow:

```yaml
env:
 MY_SECRET: ${{ secrets.MY_SECRET_NAME }}
```

## Troubleshooting

### Build Fails on Linux

**Symptom:** Missing system libraries

**Solution:** Add system dependencies:

```yaml
- name: Install system dependencies (Linux)
 if: runner.os == 'Linux'
 run: |
   sudo apt-get update
   sudo apt-get install -y libcurl4-openssl-dev libxml2-dev
```

### Build Fails: App Directory Not Found

**Symptom:** "App directory 'app' not found"

**Solution:** Ensure your Shiny app is in the correct directory and update
`APP_DIR` if needed.

### R Package Installation Fails

**Symptom:** "Package not available" errors

**Solution:** Ensure packages are available on CRAN or add custom
repositories:

```yaml
- name: Install R dependencies
 uses: r-lib/actions/setup-r-dependencies@v2
 with:
   extra-packages: |
     any::shinyelectron
     github::user/package
```

### Build Times Out

**Symptom:** Job exceeds 6 hour limit

**Solution:** Optimize your build or split into smaller jobs. Consider
building fewer platforms per workflow run.

### Artifacts Not Found

**Symptom:** Upload step shows "No files found"

**Solution:** Check the build output path. Run `ls -laR build/` to debug:

```yaml
- name: Debug build output
 run: ls -laR build/
 shell: bash
```

### Using sitrep for Diagnostics

Add diagnostic output to your workflow:

```yaml
- name: Run diagnostics
 run: |
   Rscript -e "
     library(shinyelectron)
     sitrep_shinyelectron()
   "
```

## Complete Example

Here's a minimal working example:

**Repository structure:**

```
my-app/
├── .github/workflows/build-electron.yml
├── app/
│   └── app.R
└── README.md
```

**app/app.R:**

```r
library(shiny)

ui <- fluidPage(
 titlePanel("Hello Shiny!"),
 sliderInput("n", "Number:", 1, 100, 50),
 textOutput("result")
)

server <- function(input, output) {
 output$result <- renderText(paste("You selected:", input$n))
}

shinyApp(ui, server)
```

**Workflow trigger:** Push to main or create a tag like `v1.0.0`.

## Next Steps

- **[Getting Started](getting-started.html)**: Local development workflow
- **[Configuration](configuration.html)**: Customize with `_shinyelectron.yml`
- **[Troubleshooting](troubleshooting.html)**: Diagnose build issues
