---
title: "Building with GitHub Actions"
vignette: >
  %\VignetteIndexEntry{Building with GitHub Actions}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

```{r}
#| include: false
library(shinyelectron)

# Helper: read a labelled block from the bundled workflow so the vignette
# stays in sync with what shinyelectron actually ships.
show_template_yaml <- function(anchor_re, max_lines = 200) {
  path <- system.file(
    "templates", "github-actions-build.yml",
    package = "shinyelectron"
  )
  lines <- readLines(path)
  start <- grep(anchor_re, lines)[1]
  if (is.na(start)) stop("Anchor not found: ", anchor_re)

  # Walk forward while the line is blank or starts with whitespace:
  # the YAML block ends when we hit the next top-level key.
  i <- start + 1
  end <- i
  while (i <= length(lines) && (i - start) < max_lines) {
    if (lines[i] == "" || grepl("^\\s+", lines[i]) || grepl("^#", lines[i])) {
      end <- i
      i <- i + 1
    } else {
      break
    }
  }

  cat(c("```yaml", lines[start:end], "```"), sep = "\n")
}
```

A desktop installer has to be built on the OS it targets: a `.dmg` on macOS, an `.exe` on Windows, an `.AppImage` on Linux, and once again per architecture. That is six builds for full coverage, and most teams do not have all six machines on a desk. GitHub Actions rents them by the minute, runs them in parallel, and hands back the installers as artifacts. One push, six builds, no hardware juggling.

![The build matrix: one push fans out across platform runners, each producing an installer; a tag push adds a release job that attaches them all.](../man/figures/ci-build-matrix.svg){fig-alt="A git push node on the left fans out to six runner rows (macos-latest, macos-15-intel, windows-latest, windows-11-arm, ubuntu-latest, ubuntu-24.04-arm), each producing a platform-specific installer, which fan back in to a Release job on the right that runs only on tag pushes."}

## Why automate

Doing this by hand is slow and hard to reproduce. CI fixes four specific things at once:

| Problem | What CI gives you |
|---------|-------------------|
| You need macOS, Windows, and Linux hardware | Hosted runners for each |
| Local builds drift with your laptop's state | Fresh, versioned environments every run |
| Uploading binaries to a Release page by hand | Artifacts and releases produced by a workflow step |
| Platform-specific regressions slip through | The matrix runs in parallel and surfaces them on every push |

## Before you start

You need:

1. A GitHub repo containing your Shiny app.
2. The app in a subdirectory, `app/` by default.
3. Optionally, a `_shinyelectron.yml` alongside the app.

A typical layout:

```
my-shiny-project/
├── .github/
│   └── workflows/
│       └── build-electron.yml
├── app/
│   ├── app.R
│   └── ...
├── _shinyelectron.yml
└── README.md
```

## Use the bundled workflow

shinyelectron ships a ready-to-run workflow at `inst/templates/github-actions-build.yml`. Copy it into your repo:

```{r}
#| eval: false
template <- system.file(
  "templates", "github-actions-build.yml",
  package = "shinyelectron"
)

dir.create(".github/workflows", recursive = TRUE, showWarnings = FALSE)
file.copy(
  template,
  ".github/workflows/build-electron.yml"
)
```

Or grab it directly from [GitHub](https://github.com/coatless-rpkg/shinyelectron/blob/main/inst/templates/github-actions-build.yml).

The workflow runs three jobs in sequence: a build matrix, a release step gated on tag pushes, and a summary recap.

### Configure the env vars

Edit four variables at the top of the workflow. They are the only fields most projects need to change:

```{r}
#| echo: false
#| results: asis
show_template_yaml("^env:")
```

`APP_DIR` is the path to your Shiny app inside the repo. `APP_NAME` becomes the installer's display name. `NODE_VERSION` and `R_VERSION` pin the toolchain; leave them at the defaults unless you have a reason to deviate.

### What the matrix builds

The matrix spreads installers across six runners. Each runner starts from a clean image:

| Runner | Platform | Architecture | Output |
|--------|----------|--------------|--------|
| `macos-latest` | macOS | arm64 (Apple Silicon) | `.dmg` |
| `macos-15-intel` | macOS | x64 (Intel) | `.dmg` |
| `windows-latest` | Windows | x64 | `.exe` |
| `windows-11-arm` | Windows | arm64 | `.exe` |
| `ubuntu-latest` | Ubuntu | x64 | `.AppImage` |
| `ubuntu-24.04-arm` | Ubuntu | arm64 | `.AppImage` |

CPU and RAM allocations come from GitHub's hosted-runner specs, which evolve over time; check [the GitHub-hosted runners documentation](https://docs.github.com/en/actions/using-github-hosted-runners/about-github-hosted-runners) for current numbers.

Every runner steps through the same recipe:

1. Checkout the repo.
2. Set up R with `r-lib/actions/setup-r`.
3. Set up Node.js (version from `NODE_VERSION`).
4. Install `shinyelectron` and dependencies.
5. Run `shinyelectron::export()`.
6. Upload the build output as a run artifact.

A release job runs after the matrix when the trigger is a tag push, downloading every artifact and attaching them to a fresh GitHub Release.

### Push and tag

Commit and push to fire the workflow on `main` or `master`:

```bash
git add .github/workflows/build-electron.yml
git commit -m "Add Electron build workflow"
git push
```

Tag a version to cut a release:

```bash
git tag v1.0.0
git push origin v1.0.0
```

Tags containing `-alpha` or `-beta` are marked as pre-releases automatically.

### Status badge

Drop a badge in your README so contributors see build state at a glance:

```markdown
[![Release](https://github.com/YOUR-USERNAME/YOUR-REPO/actions/workflows/build-electron.yml/badge.svg)](https://github.com/YOUR-USERNAME/YOUR-REPO/actions/workflows/build-electron.yml)
```

## Customising

The bundled workflow expects most projects to adjust four things: where the app lives, which platforms to ship, what icons to use, and whether to cache R packages aggressively. Each can be edited in place.

### App in a different folder

Override `APP_DIR`:

```yaml
env:
  APP_DIR: 'src/shiny-app'
```

### Narrower platform list

Trim the matrix to what you ship. Each entry corresponds to one runner; remove the rest:

```yaml
strategy:
  matrix:
    include:
      - os: macos-latest
        platform: mac
        arch: arm64
      - os: windows-latest
        platform: win
        arch: x64
```

### Custom icons

Ship icons from the repo and pass them to `export()`:

```yaml
- name: Build Electron app
  run: |
    Rscript -e "
      shinyelectron::export(
        appdir = '${{ env.APP_DIR }}',
        destdir = 'build',
        icon = 'assets/icon.icns'
      )
    "
```

::: {.callout-note}
Icon requirements: macOS uses `.icns` (build with `iconutil`), Windows uses `.ico` (multi-resolution recommended), Linux uses `.png` at 512x512.
:::

### Caching R packages

`npm` caching is on by default. R packages are worth caching too:

```yaml
- name: Setup R
  uses: r-lib/actions/setup-r@v2
  with:
    r-version: ${{ env.R_VERSION }}
    use-public-rspm: true

- name: Cache R packages
  uses: actions/cache@v5
  with:
    path: ${{ env.R_LIBS_USER }}
    key: ${{ runner.os }}-r-${{ hashFiles('**/DESCRIPTION') }}
```

### Config file wins, workflow overrides

A `_shinyelectron.yml` in the app directory is picked up automatically. Workflow parameters passed to `shinyelectron::export()` override its values when set.

```yaml
app:
  name: "My Shiny Dashboard"
  version: "1.0.0"

window:
  width: 1400
  height: 900

build:
  type: "r-shiny"
  runtime_strategy: "shinylive"
```

## Signing in CI

Signing keeps the same `electron-builder` environment variables as a local build, just stored as GitHub Secrets instead. Open Settings, Secrets and variables, Actions, and add each value by name. Reference them from the build job's `env`:

```yaml
- name: Build Electron app
  env:
    # macOS
    CSC_LINK: ${{ secrets.MAC_CERTIFICATE }}
    CSC_KEY_PASSWORD: ${{ secrets.MAC_CERTIFICATE_PASSWORD }}
    APPLE_ID: ${{ secrets.APPLE_ID }}
    APPLE_APP_SPECIFIC_PASSWORD: ${{ secrets.APPLE_APP_SPECIFIC_PASSWORD }}
    # Windows
    WIN_CSC_LINK: ${{ secrets.WIN_CERTIFICATE }}
    WIN_CSC_KEY_PASSWORD: ${{ secrets.WIN_CERTIFICATE_PASSWORD }}
  run: |
    Rscript -e "
      shinyelectron::export(
        appdir = '${{ env.APP_DIR }}',
        destdir = 'build',
        sign = TRUE
      )
    "
```

::: {.callout-warning}
Certificates come from Apple (macOS) and a commercial CA (Windows). Unsigned apps trigger Gatekeeper and SmartScreen warnings on end-user machines. See [Code Signing and Distribution](code-signing.html) for the full setup.
:::

## Coming soon: a composite action

A composite GitHub Action at [`coatless-actions/shiny-to-electron`](https://github.com/coatless-actions/shiny-to-electron) is in development. The goal is to wrap the whole build pipeline so a workflow shrinks to a few lines:

```yaml
on:
  push:
    tags: ['v*']
jobs:
  build:
    runs-on: ${{ matrix.os }}
    strategy:
      matrix:
        os: [ubuntu-latest, macos-latest, windows-latest]
    steps:
      - uses: actions/checkout@v6
      - uses: coatless-actions/shiny-to-electron@v1
        with:
          appdir: app
```

The action does not have a published release yet, so the bundled template above is the path to use today. Once the action ships, this section will become the recommended quickstart.

## CI-specific troubleshooting

The general guide in [Troubleshooting](troubleshooting.html) covers symptoms that show up on any machine. The items below are CI-only or turn up much more often on hosted runners than on a developer laptop.

### Linux build fails on missing libraries

Hosted Ubuntu runners are minimal. Install whatever system packages your R or Python dependencies need before the build:

```yaml
- name: Install system dependencies (Linux)
  if: runner.os == 'Linux'
  run: |
    sudo apt-get update
    sudo apt-get install -y libcurl4-openssl-dev libxml2-dev
```

### `appdir` points at the wrong directory

A workflow that hardcodes `appdir: app` will fail with `App directory 'app' not found` if your repo puts the Shiny code somewhere else. Update `APP_DIR` to match the actual path.

### R package install fails on a runner but not locally

Most often the package is not on CRAN and the workflow never told `setup-r-dependencies` where to find it. Add the repo or the GitHub source explicitly:

```yaml
- name: Install R dependencies
  uses: r-lib/actions/setup-r-dependencies@v2
  with:
    extra-packages: |
      any::shinyelectron
      github::user/package
```

### Job hits the six-hour limit

GitHub-hosted runners cap individual jobs at six hours. If a build comes close, shrink the matrix, cache packages more aggressively, or split the build into separate workflows that run in parallel.

### Run sitrep on the runner

When a build is failing in CI but works on your laptop, run the diagnostic on the runner itself to compare environments:

```yaml
- name: Run diagnostics
  run: |
    Rscript -e "
      library(shinyelectron)
      sitrep_shinyelectron()
    "
```

## Next steps

- [Getting Started](getting-started.html): local development workflow.
- [Configuration](configuration.html): customize with `_shinyelectron.yml`.
- [Troubleshooting](troubleshooting.html): diagnose build issues.
