---
title: "Node.js Management"
vignette: >
  %\VignetteIndexEntry{Node.js Management}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

```{r}
#| include: false
library(shinyelectron)
```

The shinyelectron package can manage its own Node.js installation, separate from
any system-wide installation. This guide explains how to install, manage, and
troubleshoot Node.js within shinyelectron.

## Why Local Node.js?

Managing Node.js locally through shinyelectron provides several advantages:

| Benefit | Description |
|---------|-------------|
| **No admin rights** | Install without sudo/administrator privileges |
| **Isolation** | Doesn't affect system Node.js or other projects |
| **Reproducibility** | Pin specific versions for consistent builds |
| **Portability** | Works in environments where system installation isn't possible |

## Installing Node.js

### Basic Installation

Install the latest LTS (Long Term Support) version:

```{r}
#| eval: false
install_nodejs()
```
```
ℹ Detecting latest Node.js LTS version...
✔ Latest LTS: v22.11.0
── Installing Node.js v22.11.0 ───────────────────────────────────────────────
ℹ Platform: mac
ℹ Architecture: arm64
ℹ Fetching checksums...
ℹ Downloading from https://nodejs.org/dist/v22.11.0/node-v22.11.0-darwin-arm64.tar.gz
✔ Checksum verified
ℹ Extracting archive...
✔ Node.js v22.11.0 installed successfully
ℹ Location: ~/.shinyelectron/cache/assets/nodejs/v22.11.0/darwin-arm64
ℹ shinyelectron will automatically use this installation
```

### Installing a Specific Version

Pin a specific version for reproducible builds:

```{r}
#| eval: false
install_nodejs(version = "20.18.0")
```

### Reinstalling (Force)

Force reinstall to fix corrupted installations:

```{r}
#| eval: false
install_nodejs(force = TRUE)
```

## Version Management

### Checking Installed Versions

shinyelectron reports on local Node.js installations in system diagnostics:

```{r}
#| eval: false
sitrep_electron_system()
```
```
── System Requirements Report ────────────────────────────────────────────────
✔ Platform: darwin
✔ Architecture: arm64
✔ Local Node.js (shinyelectron): v22.11.0 ✓
  Other versions: 20.18.0
✔ Active Node.js: v22.11.0 (local) ✓
✔ npm: v10.9.0 ✓
```

### Multiple Versions

You can have multiple Node.js versions installed simultaneously:

```{r}
#| eval: false
# Install multiple versions
install_nodejs(version = "22.11.0")
install_nodejs(version = "20.18.0")
install_nodejs(version = "18.20.0")
```

The most recently installed version (or the latest version if versions are
equal) is used by default.

## How It Works

### Installation Location

Node.js installations are stored in your user cache directory:

```
~/.shinyelectron/cache/assets/nodejs/
├── v22.11.0/
│   └── darwin-arm64/      # macOS ARM (M1/M2/M3)
│       ├── bin/
│       │   ├── node       # Node.js executable
│       │   ├── npm        # npm executable
│       │   └── npx        # npx executable
│       ├── include/
│       └── lib/
├── v20.18.0/
│   ├── darwin-arm64/
│   └── win-x64/           # Windows 64-bit
│       ├── node.exe
│       ├── npm.cmd
│       └── npx.cmd
└── ...
```

### Checksum Verification

Downloads are verified using SHA256 checksums from the official Node.js
distribution:

1. shinyelectron downloads `SHASUMS256.txt` from nodejs.org
2. Downloads the Node.js archive
3. Computes the SHA256 hash of the downloaded file
4. Compares against the expected checksum
5. Aborts if checksums don't match

This ensures the downloaded files haven't been corrupted or tampered with.

### Platform Detection

shinyelectron automatically detects your platform and architecture:

| Platform | Detection | Node.js Platform |
|----------|-----------|------------------|
| macOS | `Sys.info()[["sysname"]] == "Darwin"` | darwin |
| Windows | `Sys.info()[["sysname"]] == "Windows"` | win |
| Linux | `Sys.info()[["sysname"]] == "Linux"` | linux |

| Architecture | Detection | Node.js Arch |
|--------------|-----------|--------------|
| x86-64 | Intel/AMD 64-bit | x64 |
| ARM64 | Apple Silicon, ARM servers | arm64 |

## Auto-Installation

### Via Configuration File

Enable automatic installation in `_shinyelectron.yml`:

```yaml
nodejs:
  auto_install: true
```

With this setting, `export()` will automatically install Node.js if not found.

### Via Configuration File with Specific Version

Pin a specific version for auto-installation:

```yaml
nodejs:
  version: "22.11.0"
  auto_install: true
```

::: {.callout-note}
## Opt-In Only

Auto-installation is disabled by default. You must explicitly set
`auto_install: true` or call `install_nodejs()` manually.
:::

## Priority and Fallback

When shinyelectron needs Node.js, it checks in this order:

1. **Local installation** (shinyelectron-managed)
2. **System installation** (if local not found)

This means your local shinyelectron installation takes priority over any
system-wide Node.js.

To use the system Node.js instead of local (not recommended):

```{r}
#| eval: false
# Internal function - not typically needed
get_node_command(prefer_local = FALSE)
```

## Clearing the Cache

### Remove All Node.js Installations

```{r}
#| eval: false
cache_clear("nodejs")
```
```
✔ Cleared nodejs cache
```

### Remove All Cached Assets

```{r}
#| eval: false
cache_clear("all")
```

This removes:

- Node.js installations
- npm packages cache
- R-related caches

## Troubleshooting

### Installation Fails: Network Error

**Symptom:**
```
✖ Failed to download Node.js
x URL: https://nodejs.org/dist/v22.11.0/node-v22.11.0-darwin-arm64.tar.gz
x Error: could not resolve host
```

**Solutions:**

1. Check your internet connection
2. Try again later (nodejs.org may be temporarily unavailable)
3. Check proxy settings if behind a corporate firewall

### Installation Fails: Checksum Mismatch

**Symptom:**
```
✖ Checksum verification failed
x Downloaded file may be corrupted
```

**Solutions:**

```{r}
#| eval: false
# Retry with force to re-download
install_nodejs(force = TRUE)
```

If the problem persists, nodejs.org may be having issues. Wait and try again
later.

### Wrong Architecture

**Symptom:** Node.js runs but Electron fails with architecture mismatch errors.

**Solution:**

Check what architecture was installed:

```{r}
#| eval: false
sitrep_electron_system()
```

If the architecture is wrong, clear and reinstall:

```{r}
#| eval: false
cache_clear("nodejs")
install_nodejs()
```

### Local Node.js Not Being Used

**Symptom:** System Node.js is used instead of local installation.

**Cause:** The local installation may be incomplete or corrupted.

**Solution:**

```{r}
#| eval: false
# Clear and reinstall
cache_clear("nodejs")
install_nodejs()

# Verify
sitrep_electron_system()
```

## Comparison with System Installation

| Aspect | Local (shinyelectron) | System |
|--------|----------------------|--------|
| Installation | `install_nodejs()` | Download from nodejs.org |
| Location | `~/.shinyelectron/cache/` | `/usr/local/`, `C:\Program Files\` |
| Permissions | User-only | Often requires admin |
| Updates | Manual via `install_nodejs()` | System package manager |
| Isolation | Per-user, isolated | System-wide, shared |
| Version control | Multiple versions supported | Usually one version |

For most shinyelectron users, the local installation is recommended because it's
isolated and reproducible.

## Next Steps

- **[Getting Started](getting-started.html)**: First-time user walkthrough
- **[Configuration](configuration.html)**: Set Node.js options in config files
- **[Troubleshooting](troubleshooting.html)**: Diagnose other issues
