---
title: "Troubleshooting Guide"
vignette: >
  %\VignetteIndexEntry{Troubleshooting Guide}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

```{r}
#| include: false
library(shinyelectron)
```

When something breaks, start with a diagnostic. shinyelectron's `sitrep_*`
functions (short for "situation report") inspect your environment and tell
you what is and is not ready. This guide covers the diagnostics, then
walks through fixes, grouped by where the problem shows up: install,
build, or runtime.

## Start with a full report

When you hit a problem, run the full diagnostic first.

```{r}
#| eval: false
sitrep_shinyelectron()
```
```
-- Complete shinyelectron Diagnostic Report ----------------------------------
-------------------------------------------------------------------------------

-- System Requirements Report -----------------------------------------------
v Platform: darwin
v Architecture: arm64
v Local Node.js (shinyelectron): v22.11.0
v Active Node.js: v22.11.0 (local)
v npm: v10.9.0
v R: v4.4.2

-- Python
v Python 3.12.4
v Python shinylive: 0.7.0 (ready)
v Python shiny: 1.2.1 (ready)

-- Container Engine
v Container engine: docker

-- Cached Runtimes
i No cached runtimes found

v All system requirements satisfied
-------------------------------------------------------------------------------

-- Dependencies Report ------------------------------------------------------
-- Required Packages
v cli: v3.6.5
v fs: v1.6.6
v jsonlite: v2.0.0
v processx: v3.8.6
v whisker: v0.4.1
v utils: v4.4.2
v tools: v4.4.2
v All required dependencies satisfied
-------------------------------------------------------------------------------

-- Build Tools Report -------------------------------------------------------
i Checking build tools for platform: darwin
v Xcode Command Line Tools: Found
v Build tools ready
-------------------------------------------------------------------------------

-- Overall Summary ----------------------------------------------------------
v All systems ready! You should be able to build Electron apps successfully
```

This runs every check and ranks any issues by priority. The system report
covers Python availability, the `shinylive` and `shiny` Python packages,
the container engine, and any cached runtimes.

## Targeted diagnostics

Each area has its own function. Run one when you already know which
subsystem is unhappy.

| Function | What it checks |
|----------|----------------|
| `sitrep_electron_system()` | Node.js, npm, platform, R, Python (plus `shinylive` and `shiny`), container engine, cached runtimes |
| `sitrep_electron_dependencies()` | Required and optional R packages |
| `sitrep_electron_build_tools()` | Platform-specific build tools |
| `sitrep_electron_project()` | Electron project structure |

### System requirements

```{r}
#| eval: false
sitrep_electron_system()
```

Covers:

- **Platform and architecture** (darwin, win, linux; x64 or arm64).
- **Local Node.js** installed by shinyelectron.
- **Active Node.js** version, whether local or system.
- **npm** version.
- **R** version.
- **Python** plus the `shinylive` and `shiny` Python packages.
- **Container engine** (Docker or Podman).
- **Cached runtimes** previously downloaded for R or Python.

::: {.callout-tip}
## Local vs system Node.js

shinyelectron prefers its own Node.js install when one exists. See
[Node.js Management](nodejs-management.html) for the full resolution
order.
:::

### R package dependencies

```{r}
#| eval: false
sitrep_electron_dependencies()
```

Required:

- `cli`, formatted console output
- `fs`, file system operations
- `jsonlite`, JSON parsing
- `processx`, process execution
- `whisker`, template rendering
- `utils`, utility functions
- `tools`, R tools

Optional:

- `shinylive`, required for R apps built with the `shinylive` runtime strategy
- `DT`, enhanced data tables (for your apps)
- `ggplot2`, graphics (for your apps)

### Build tools

```{r}
#| eval: false
sitrep_electron_build_tools()
```

| Platform | Required tools |
|----------|----------------|
| macOS | Xcode Command Line Tools |
| Windows | Visual Studio Build Tools, Python |
| Linux | gcc, make (build-essential) |

### Electron project

```{r}
#| eval: false
# Check current directory
sitrep_electron_project()

# Check a specific directory
sitrep_electron_project("path/to/electron-app")
```

Validates:

- `package.json` existence and contents
- `main.js` presence
- App files in expected locations
- Build scripts completeness
- `node_modules` installation status

## Install-time problems

These are things that break before you run a build: missing runtimes,
missing R packages, missing compilers.

### Node.js not found

```
x Node.js: Not found
```

Pick one:

1. Install locally with shinyelectron (recommended).

   ```{r}
   #| eval: false
   install_nodejs()
   ```

2. Install from <https://nodejs.org/>.

3. Enable auto-install in config.

   ```yaml
   nodejs:
     auto_install: true
   ```

### Node.js version too old

```
! Node.js: v18.20.0 (version 22+ required)
```

shinyelectron requires Node.js 22.0.0 or newer. Older LTS lines (18.x,
20.x) miss features the bundled Electron version needs.

Install a newer local copy.

```{r}
#| eval: false
# Install latest LTS (currently 22.x)
install_nodejs()

# Or pin a specific version
install_nodejs(version = "22.11.0")
```

### npm not found

```
x npm: Not found
```

npm ships with Node.js. If Node is installed but npm is gone, reinstall
Node.

```{r}
#| eval: false
install_nodejs(force = TRUE)
```

### Missing R packages

```
x shinylive: Not installed
```

Install the one that is missing.

```{r}
#| eval: false
install.packages("shinylive")
```

Or install several at once.

```{r}
#| eval: false
install.packages(c("shinylive", "cli", "fs"))
```

### Xcode Command Line Tools missing (macOS)

```
! Xcode Command Line Tools: Not found
```

In Terminal:

```bash
xcode-select --install
```

### Visual Studio Build Tools missing (Windows)

```
! Visual Studio Build Tools: Not found
```

1. Download from <https://visualstudio.microsoft.com/downloads/>.
2. Install the "Desktop development with C++" workload.
3. Restart your R session.

### Build tools missing (Linux)

```
! Build tools: Incomplete
```

Ubuntu or Debian:

```bash
sudo apt-get update
sudo apt-get install build-essential
```

Fedora or RHEL:

```bash
sudo dnf groupinstall "Development Tools"
```

### Python not found

```
i Python: not found (needed for py-shiny apps)
```

1. Install Python from <https://www.python.org/downloads/>. Version 3.9
   or newer.

2. On **Windows**, tick "Add python.exe to PATH" during install. If you
   skipped that, add the install directory manually. It usually lives at
   `C:\Users\<you>\AppData\Local\Programs\Python\Python3XX\`.

3. On **macOS**, the `python3` that ships with Xcode Command Line Tools
   usually works. If it does not, use Homebrew: `brew install python`.

4. Verify Python is reachable.

   ```{r}
   #| eval: false
   Sys.which("python3")
   # On Windows, Sys.which("python") is also checked
   ```

### Python shinylive CLI not working

```
x The shinylive Python package CLI is required for the shinylive strategy with Python apps
  ... cannot be directly executed
  ... No module named shinylive.__main__
```

The `shinylive` package ships a console script, not a `__main__.py`, so
`python -m shinylive` fails. The real problem is that the console script
is not on PATH.

1. Reinstall to recreate the script.

   ```bash
   pip install --upgrade --force-reinstall shinylive
   ```

2. On **Windows**, pip installs scripts in
   `%APPDATA%\Python\Python3XX\Scripts`. Add that directory to PATH.

   ```powershell
   # Find where pip installed the script
   python -c "import sysconfig; print(sysconfig.get_path('scripts'))"

   # Add it to your user PATH
   $scriptsDir = python -c "import sysconfig; print(sysconfig.get_path('scripts'))"
   [Environment]::SetEnvironmentVariable('Path', "$env:Path;$scriptsDir", 'User')
   ```

   Restart your R session.

3. Verify.

   ```bash
   shinylive --version
   ```

### Python shiny package not installed

```
i Python shiny: not usable
```

```bash
pip install shiny
```

This is needed when `py-shiny` runs under any non-shinylive strategy
(`system`, `bundled`, `auto-download`, `container`). The `shinylive`
strategy uses the `shinylive` Python package instead.

### Docker or Podman not found

```
i Docker/Podman: not found (needed for container strategy)
```

Install one:

- Docker Desktop: <https://docs.docker.com/get-docker/>
- Podman: <https://podman.io/getting-started/installation>

Confirm the engine is running.

```bash
docker info
# or
podman info
```

::: {.callout-warning}
## Docker on Parallels / nested virtualization

Docker Desktop needs hardware virtualization. If you run macOS inside a
Parallels VM (common on Apple Silicon), Docker Desktop will fail because
Parallels does not expose nested virtualization to the guest.

Workaround: use the `system`, `bundled`, or `auto-download` runtime
strategy, or run the Docker build on the host.
:::

## Build-time problems

These appear during `export()`, `npm install`, or the electron-builder
packaging step.

### package.json not found

```
x package.json: Not found
```

You pointed the diagnostic at the wrong directory. Point it at the
Electron sub-project.

```{r}
#| eval: false
sitrep_electron_project("my-export/electron-app")
```

### node_modules not found

```
i node_modules: Not found (run 'npm install')
```

Install the project's dependencies.

```{r}
#| eval: false
setwd("my-export/electron-app")
system("npm install")
```

Or from a terminal:

```bash
cd my-export/electron-app
npm install
```

### Container image pull failure

```
Error: ... manifest unknown ... or image not found
```

1. Check the image name in `_shinyelectron.yml`.

   ```yaml
   container:
     image: "shinyelectron/r-shiny"
     tag: "latest"
   ```

2. Pull the image manually to rule out a network or auth problem.

   ```bash
   docker pull shinyelectron/r-shiny:latest
   ```

3. For private registries, run `docker login` first.

### Container port conflict

```
Error: ... bind: address already in use
```

Something else is on the port already, often a container you forgot to
stop.

1. Move to a different port in `_shinyelectron.yml`.

   ```yaml
   server:
     port: 3839
   ```

2. Or stop the other container.

   ```bash
   docker ps                     # find the running container
   docker stop <container-id>    # stop it
   ```

### GNU tar vs bsdtar on Windows

A runtime archive fails to extract with:

```
Cannot connect to C: resolve failed
```

Git for Windows puts GNU tar on the PATH. GNU tar reads `C:\...` as a
remote host spec (`host:path`), not a drive letter. The runtime
downloader needs Windows 10's bundled bsdtar at
`%SystemRoot%\System32\tar.exe`, which understands drive letters.

shinyelectron already handles this: the JavaScript downloader resolves
the full path to `System32\tar.exe`, and the R-side extraction uses R's
internal tar. If you still hit the error, check what is actually on PATH.

```bash
where tar
```

The first hit should be `C:\Windows\System32\tar.exe`, not
`C:\Program Files\Git\usr\bin\tar.exe`.

### EBUSY file locks when rebuilding (Windows)

```
Error: EBUSY: resource busy or locked, unlink '...\electron.exe'
```

A running Electron process on Windows holds a lock on its own executable
and DLLs. A rebuild cannot overwrite the locked files.

1. Close the running Electron app.
2. If the process lingers, kill `electron.exe` from Task Manager.
3. Re-run the build.

### Slow builds (Windows)

Symptom: builds crawl, especially the `npm install` step and final
packaging.

Cause: Windows Defender real-time scanning inspects every file `npm`
writes inside `node_modules`. That is thousands of small files.

1. Add a Defender exclusion for the project directory.

   Settings > Privacy & Security > Virus & threat protection > Manage
   settings > Exclusions > Add an exclusion > Folder.

2. Add an exclusion for the shinyelectron cache at `~/.shinyelectron`.

## Runtime problems

These show up after the app is packaged, when the end user (or you)
launches it.

### R not on PATH (Windows)

Symptom: a packaged app using the `system` runtime cannot find `Rscript`
on the user's machine.

Cause: the default R installer on Windows puts R under `C:\Program Files`
(or `C:\Program Files (x86)` for 32-bit) and does not add it to PATH.

1. During R installation, tick the option to add R to PATH.
2. After installation, add the R `bin` directory manually, for example:

   ```
   C:\Program Files\R\R-4.4.2\bin
   ```

3. For production, use the `bundled` or `auto-download` runtime strategy
   so end users do not need R pre-installed.

### GateKeeper blocks the app (macOS)

Symptom: users see "app is damaged and can't be opened" or "app can't be
opened because Apple cannot check it for malicious software".

Cause: GateKeeper blocks unsigned or unnotarized apps.

For **distribution**, sign and notarize. Configure in
`_shinyelectron.yml`:

```yaml
signing:
  sign: true
  mac:
    identity: "Developer ID Application: Your Name (TEAMID)"
    team_id: "TEAMID"
    notarize: true
```

Set `APPLE_TEAM_ID`, `APPLE_ID`, and `APPLE_APP_SPECIFIC_PASSWORD` before
building.

For **local testing**, right-click the app and choose Open, then confirm.
That bypasses GateKeeper for that app only. Or strip the quarantine
attribute:

```bash
xattr -cr /Applications/MyApp.app
```

### SmartScreen blocks the installer (Windows)

Symptom: SmartScreen shows "Windows protected your PC" when the installer
or app is run.

Cause: the executable is unsigned or lacks SmartScreen reputation.

For **distribution**, sign the app with a code signing certificate.

```yaml
signing:
  sign: true
  win:
    certificate_file: "path/to/cert.pfx"
```

Set `CSC_LINK` and `CSC_KEY_PASSWORD` before building.

For **local testing**, click "More info" then "Run anyway".

EV certificates receive SmartScreen reputation immediately. Standard
certificates build reputation over time, as more users run the signed
installer.

## Reading sitrep results in code

Every `sitrep_*` function returns its results invisibly. Capture them
when you want to automate checks.

```{r}
#| eval: false
# Get results without printing
results <- sitrep_shinyelectron(verbose = FALSE)

# Check for issues
if (length(results$system$issues) > 0) {
  cat("System issues found:\n")
  print(results$system$issues)
}

# Check Node.js status
if (results$system$node$installed) {
  cat("Node.js version:", results$system$node$version, "\n")
}
```

## Clear the cache

When nothing else explains a problem, clear the cache.

```{r}
#| eval: false
# Clear all cached assets
cache_clear("all")

# Clear only Node.js installs
cache_clear("nodejs")

# Clear the npm cache
cache_clear("npm")
```

Reinstall Node.js after a clear.

```{r}
#| eval: false
install_nodejs()
```

## Debugging a packaged app

A packaged Electron app is quiet by default: only warnings and errors
print. To watch runtime detection, dependency installation, and server
startup, launch the app with `SHINYELECTRON_DEBUG=1`.

```bash
# macOS / Linux
SHINYELECTRON_DEBUG=1 /path/to/YourApp.app/Contents/MacOS/YourApp

# Windows (PowerShell)
$env:SHINYELECTRON_DEBUG = "1"; & "C:\Path\To\YourApp.exe"
```

Debug messages are prefixed with `[shinyelectron]` and include which
runtime was found, the spawn command for R or Python, port retries, and
server-ready events. Include this output when you file a bug report.

## Migrating from the old API

Earlier releases treated shinylive as an app type. `app_type` accepted
`r-shinylive` and `py-shinylive` alongside `r-shiny` and `py-shiny`.
That is gone. shinylive is now a runtime strategy.

| Old API | New API |
|---------|---------|
| `app_type = "r-shinylive"` | `app_type = "r-shiny"`, `runtime_strategy = "shinylive"` |
| `app_type = "py-shinylive"` | `app_type = "py-shiny"`, `runtime_strategy = "shinylive"` |
| `app_type = "r-shiny"` | `app_type = "r-shiny"` (pick any `runtime_strategy`) |
| `app_type = "py-shiny"` | `app_type = "py-shiny"` (pick any `runtime_strategy`) |

`app_type` is now optional. shinyelectron autodetects `r-shiny` or
`py-shiny` from the files in `appdir`. `runtime_strategy` defaults to
`"shinylive"` when neither the function arg nor the config sets one.

The legacy values still work, but they emit a deprecation warning. If
you see a warning about `r-shinylive` or `py-shinylive`, rewrite the
call to the new two-axis form. Update `_shinyelectron.yml` the same way:
replace `build.type: "r-shinylive"` with either
`build.runtime_strategy: "shinylive"` (relying on autodetect) or
`build.type: "r-shiny"` plus `build.runtime_strategy: "shinylive"`
(explicit).

## Getting help

If every diagnostic is green but something still breaks:

1. Read the console output for the exact error message. Enable
   `SHINYELECTRON_DEBUG=1` for verbose backend logs.
2. Confirm your Shiny app runs locally with `shiny::runApp()`.
3. Check file permissions on your project directory.
4. Open an issue at
   <https://github.com/coatless-rpkg/shinyelectron/issues>.

When you file an issue, include the output of:

```{r}
#| eval: false
sitrep_shinyelectron()
```

## Next steps

- **[Getting Started](getting-started.html)**: First-time user walkthrough
- **[Configuration](configuration.html)**: Customize your builds
- **[Node.js Management](nodejs-management.html)**: Manage Node.js installs
