---
title: "Troubleshooting Guide"
vignette: >
  %\VignetteIndexEntry{Troubleshooting Guide}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

```{r}
#| include: false
library(shinyelectron)
```

When something goes wrong with shinyelectron, the diagnostic functions (called
"sitrep" for "situation report") help identify and resolve issues. This guide
explains how to use these functions effectively.

## Quick Start: Run Full Diagnostics

When you encounter problems, start with a complete diagnostic check:

```{r}
#| eval: false
sitrep_shinyelectron()
```
```
── Complete shinyelectron Diagnostic Report ──────────────────────────────────
────────────────────────────────────────────────────────────────────────────────

── System Requirements Report ────────────────────────────────────────────────
✔ Platform: darwin
✔ Architecture: arm64
✔ Local Node.js (shinyelectron): v22.11.0 ✓
✔ Active Node.js: v22.11.0 (local) ✓
✔ npm: v10.9.0 ✓
✔ R: v4.4.2 ✓

── Python
✔ Python 3.12.4
✔ Python shinylive: 0.7.0 (ready)
✔ Python shiny: 1.2.1 (ready)

── Container Engine
✔ Container engine: docker

── Cached Runtimes
ℹ No cached runtimes found

✔ All system requirements satisfied ✓
────────────────────────────────────────────────────────────────────────────────

── Dependencies Report ───────────────────────────────────────────────────────
── Required Packages
✔ cli: v3.6.5 ✓
✔ fs: v1.6.6 ✓
✔ jsonlite: v2.0.0 ✓
✔ processx: v3.8.6 ✓
✔ whisker: v0.4.1 ✓
✔ utils: v4.4.2 ✓
✔ tools: v4.4.2 ✓
✔ All required dependencies satisfied ✓
────────────────────────────────────────────────────────────────────────────────

── Build Tools Report ────────────────────────────────────────────────────────
ℹ Checking build tools for platform: darwin
✔ Xcode Command Line Tools: Found ✓
✔ Build tools ready ✓
────────────────────────────────────────────────────────────────────────────────

── Overall Summary ───────────────────────────────────────────────────────────
✔ All systems ready! You should be able to build Electron apps successfully ✓
```

This command runs all diagnostic checks and provides prioritized recommendations
if issues are found. The system report now includes Python availability,
Python package checks (shinylive and shiny), container engine detection, and
cached runtime inventories.

## Individual Diagnostic Functions

Each diagnostic function focuses on a specific area:

| Function | What It Checks |
|----------|----------------|
| `sitrep_electron_system()` | Node.js, npm, platform, R version, Python (+ shinylive/shiny packages), container engine, cached runtimes |
| `sitrep_electron_dependencies()` | Required and optional R packages |
| `sitrep_electron_build_tools()` | Platform-specific build tools |
| `sitrep_electron_project()` | Electron project structure |

### System Requirements

Check Node.js, npm, and platform configuration:

```{r}
#| eval: false
sitrep_electron_system()
```

This function checks:

- **Platform and architecture** (darwin/win/linux, x64/arm64)
- **Local Node.js** installed by shinyelectron
- **Active Node.js** version (local or system)
- **npm** availability and version
- **R version** compatibility
- **Python** availability, plus whether `shinylive` and `shiny` Python packages are usable
- **Container engine** (Docker or Podman)
- **Cached runtimes** (previously downloaded R and Python runtimes)

::: {.callout-tip}
## Local vs System Node.js

shinyelectron can use either a locally installed Node.js (via `install_nodejs()`)
or your system's Node.js. The local installation is preferred when available.
:::

### R Package Dependencies

Check required and optional R packages:

```{r}
#| eval: false
sitrep_electron_dependencies()
```

**Required packages:**

- `cli` - Formatted console output
- `fs` - File system operations
- `jsonlite` - JSON parsing
- `processx` - Process execution
- `whisker` - Template rendering
- `utils` - Utility functions
- `tools` - R tools

**Optional packages:**

- `shinylive` - Required for r-shinylive app type
- `DT` - Enhanced data tables (for your apps)
- `ggplot2` - Graphics (for your apps)

### Build Tools

Check platform-specific build tools:

```{r}
#| eval: false
sitrep_electron_build_tools()
```

**Requirements by platform:**

| Platform | Required Tools |
|----------|----------------|
| macOS | Xcode Command Line Tools |
| Windows | Visual Studio Build Tools, Python |
| Linux | gcc, make (build-essential) |

### Electron Project

Check an existing Electron project structure:

```{r}
#| eval: false
# Check current directory
sitrep_electron_project()

# Check specific directory
sitrep_electron_project("path/to/electron-app")
```

This function validates:

- **package.json** existence and validity
- **main.js** presence
- **App files** in expected locations
- **Build scripts** completeness
- **node_modules** installation status

## Common Issues and Solutions

### Node.js Not Found

**Symptom:**
```
✖ Node.js: Not found
```

**Solutions:**

1. **Install locally with shinyelectron** (recommended):
   ```{r}
   #| eval: false
   install_nodejs()
   ```

2. **Install from nodejs.org:**
   Download and install from <https://nodejs.org/>

3. **Enable auto-install in config:**
   ```yaml
   nodejs:
     auto_install: true
   ```

### Node.js Version Too Old

**Symptom:**
```
⚠ Node.js: v18.20.0 (version 22+ required)
```

shinyelectron requires Node.js >= 22.0.0. Older LTS lines (18.x, 20.x) are
not supported because the Electron version used by the package depends on
features introduced in Node.js 22.

**Solution:**

Install a newer version locally:

```{r}
#| eval: false
# Install latest LTS (currently 22.x)
install_nodejs()

# Or install specific version
install_nodejs(version = "22.11.0")
```

### npm Not Found

**Symptom:**
```
✖ npm: Not found
```

**Solution:**

npm is bundled with Node.js. If Node.js is installed but npm is missing,
reinstall Node.js:

```{r}
#| eval: false
install_nodejs(force = TRUE)
```

### Missing R Packages

**Symptom:**
```
✖ shinylive: Not installed
```

**Solution:**

```{r}
#| eval: false
install.packages("shinylive")
```

For multiple missing packages:

```{r}
#| eval: false
install.packages(c("shinylive", "cli", "fs"))
```

### Xcode Command Line Tools Not Found (macOS)

**Symptom:**
```
⚠ Xcode Command Line Tools: Not found
```

**Solution:**

Run in Terminal:

```bash
xcode-select --install
```

### Visual Studio Build Tools Not Found (Windows)

**Symptom:**
```
⚠ Visual Studio Build Tools: Not found
```

**Solution:**

1. Download Visual Studio Build Tools from
   <https://visualstudio.microsoft.com/downloads/>
2. Install the "Desktop development with C++" workload
3. Restart your R session

### Build Tools Not Found (Linux)

**Symptom:**
```
⚠ Build tools: Incomplete
```

**Solution:**

Install build-essential (Ubuntu/Debian):

```bash
sudo apt-get update
sudo apt-get install build-essential
```

Install development tools (Fedora/RHEL):

```bash
sudo dnf groupinstall "Development Tools"
```

### package.json Not Found

**Symptom:**
```
✖ package.json: Not found
```

**Cause:** You're checking a directory that isn't an Electron project.

**Solution:**

Run the diagnostic on the correct directory:

```{r}
#| eval: false
# Check the electron-app subdirectory
sitrep_electron_project("my-export/electron-app")
```

### node_modules Not Found

**Symptom:**
```
ℹ node_modules: Not found (run 'npm install')
```

**Solution:**

Navigate to the Electron project and install dependencies:

```{r}
#| eval: false
# From R
setwd("my-export/electron-app")
system("npm install")
```

Or from the terminal:

```bash
cd my-export/electron-app
npm install
```

## Python Issues

### Python Not Found

**Symptom:**
```
ℹ Python: not found (needed for py-shinylive and py-shiny)
```

**Solutions:**

1. **Install Python** from <https://www.python.org/downloads/>. Version 3.9
   or newer is recommended.

2. **Windows:** During installation, check **"Add python.exe to PATH"**. If
   you already installed Python without that option, add the install directory
   manually. The default location is typically
   `C:\Users\<you>\AppData\Local\Programs\Python\Python3XX\`.

3. **macOS:** The system `python3` from Xcode Command Line Tools usually
   works. If not, install via Homebrew: `brew install python`.

4. **Verify** Python is reachable:

   ```{r}
   #| eval: false
   Sys.which("python3")
   # On Windows, Sys.which("python") is also checked
   ```

### Python shinylive CLI Not Working

**Symptom:**
```
✖ The shinylive Python package CLI is required for py-shinylive conversion
  … cannot be directly executed
  … No module named shinylive.__main__
```

The `shinylive` Python package installs a console script, but the package
itself ships no `__main__.py`, so `python -m shinylive` alone fails. This
usually means the `shinylive` console script is not on your PATH.

**Solutions:**

1. **Reinstall** to make sure the console script is created:

   ```bash
   pip install --upgrade --force-reinstall shinylive
   ```

2. **Windows PATH issue:** `pip install` places scripts in
   `%APPDATA%\Python\Python3XX\Scripts`. Add that directory to your PATH:

   ```powershell
   # Find where pip installed the script
   python -c "import sysconfig; print(sysconfig.get_path('scripts'))"

   # Add that directory to your user PATH
   $scriptsDir = python -c "import sysconfig; print(sysconfig.get_path('scripts'))"
   [Environment]::SetEnvironmentVariable('Path', "$env:Path;$scriptsDir", 'User')
   ```

   Then restart your R session.

3. **Verify** the CLI works:

   ```bash
   shinylive --version
   ```

### Python shiny Package Not Installed

**Symptom:**
```
ℹ Python shiny: not usable
```

**Solution:**

```bash
pip install shiny
```

This is only required for `py-shiny` (native Python Shiny) apps. The
`py-shinylive` app type uses the `shinylive` package instead.

## Container Issues (Docker / Podman)

### Docker or Podman Not Found

**Symptom:**
```
ℹ Docker/Podman: not found (needed for container strategy)
```

**Solution:**

Install a container engine:

- **Docker Desktop:** <https://docs.docker.com/get-docker/>
- **Podman:** <https://podman.io/getting-started/installation>

After installation, verify the engine is running:

```bash
docker info
# or
podman info
```

::: {.callout-warning}
## Docker on Parallels / nested virtualization

Docker Desktop requires hardware virtualization. If you are running macOS
inside a Parallels VM (e.g. on Apple Silicon), Docker Desktop will fail
because Parallels does not expose nested virtualization to the guest.

**Workaround:** Use the `system`, `bundled`, or `auto-download` runtime
strategy instead, or run the Docker build on the host machine.
:::

### Container Image Pull Failures

**Symptom:**
```
Error: … manifest unknown … or image not found
```

**Solutions:**

1. **Check the image name** in your `_shinyelectron.yml`:

   ```yaml
   container:
     image: "shinyelectron/r-shiny"
     tag: "latest"
   ```

2. **Network/auth issues:** Make sure you can pull manually:

   ```bash
   docker pull shinyelectron/r-shiny:latest
   ```

3. **Private registries:** Log in first with `docker login`.

### Container Port Conflicts

**Symptom:**
```
Error: … bind: address already in use
```

Another process (or a previous container) is already using the same port.

**Solutions:**

1. **Change the port** in `_shinyelectron.yml`:

   ```yaml
   server:
     port: 3839
   ```

2. **Stop the conflicting container:**

   ```bash
   docker ps                     # find the running container
   docker stop <container-id>    # stop it
   ```

## Windows-Specific Issues

### R Not on PATH

**Symptom:** The packaged app cannot find `Rscript` on the end user's
machine when using the `system` runtime strategy.

**Cause:** The default R installer on Windows places R under
`C:\Program Files` (or `C:\Program Files (x86)` for 32-bit), but does
not add it to the system PATH.

**Solutions:**

1. **During R installation:** Check the option to add R to the PATH.

2. **After installation:** Add the R `bin` directory manually. For
   example:

   ```
   C:\Program Files\R\R-4.4.2\bin
   ```

3. **In production:** Use the `bundled` or `auto-download` runtime
   strategy so that end users do not need R pre-installed.

### GNU tar vs bsdtar Conflict

**Symptom:** Extracting a runtime archive fails with:
```
Cannot connect to C: resolve failed
```

**Cause:** Git for Windows ships GNU tar on the PATH. GNU tar
interprets paths like `C:\...` as a remote host specification
(`host:path`). The runtime downloader needs Windows 10's built-in bsdtar
(`%SystemRoot%\System32\tar.exe`), which handles drive letters correctly.

shinyelectron works around this automatically: the JavaScript runtime
downloader resolves the full path to `System32\tar.exe`, and the R-side
extraction uses R's internal tar. If you still hit this error, check
whether something has overridden the tar command:

```bash
where tar
```

The first result should be `C:\Windows\System32\tar.exe`, not
`C:\Program Files\Git\usr\bin\tar.exe`.

### EBUSY File Locks When Rebuilding

**Symptom:**
```
Error: EBUSY: resource busy or locked, unlink '...\electron.exe'
```

**Cause:** On Windows, a running Electron process locks its own
executable and DLLs. If you try to rebuild the app while the previous
build is still running, `npm` / `electron-builder` cannot overwrite the
locked files.

**Solution:**

1. Close the running Electron app.
2. If the process lingers, end it from Task Manager (`electron.exe`).
3. Re-run the build.

### Windows Defender Scanning Delays

**Symptom:** Builds are extremely slow, especially the `npm install`
step and the final packaging step.

**Cause:** Windows Defender real-time scanning inspects every file that
`npm` creates inside `node_modules` (thousands of small files).

**Solutions:**

1. **Add an exclusion** for the project directory:

   Settings > Privacy & Security > Virus & threat protection >
   Manage settings > Exclusions > Add an exclusion > Folder

2. **Add an exclusion** for the shinyelectron cache directory
   (`~/.shinyelectron`).

## Code Signing

### Unsigned App Warnings on macOS (GateKeeper)

**Symptom:** Users see "app is damaged and can't be opened" or
"app can't be opened because Apple cannot check it for malicious
software" when launching the packaged app.

**Cause:** macOS GateKeeper blocks unsigned or unnotarized applications.

**Solutions:**

1. **For distribution:** Sign and notarize the app. Configure signing
   in `_shinyelectron.yml`:

   ```yaml
   signing:
     sign: true
     mac:
       identity: "Developer ID Application: Your Name (TEAMID)"
       team_id: "TEAMID"
       notarize: true
   ```

   Set the environment variables `APPLE_TEAM_ID`, `APPLE_ID`, and
   `APPLE_APP_SPECIFIC_PASSWORD` before building.

2. **For local testing only:** Right-click the app and choose "Open",
   then confirm. This bypasses GateKeeper for that specific app. Or
   remove the quarantine attribute:

   ```bash
   xattr -cr /Applications/MyApp.app
   ```

### SmartScreen Warning on Windows

**Symptom:** Windows SmartScreen shows "Windows protected your PC" when
the user tries to run the installer or app.

**Cause:** The executable is unsigned or does not have enough
SmartScreen reputation.

**Solutions:**

1. **For distribution:** Sign the app with a code signing certificate.
   Configure in `_shinyelectron.yml`:

   ```yaml
   signing:
     sign: true
     win:
       certificate_file: "path/to/cert.pfx"
   ```

   Set the `CSC_LINK` and `CSC_KEY_PASSWORD` environment variables
   before building.

2. **For local testing:** Click "More info" then "Run anyway".

3. **EV certificates** receive immediate SmartScreen reputation.
   Standard certificates build reputation over time as more users
   run the signed installer.

## Accessing Diagnostic Results Programmatically

All sitrep functions return their results invisibly for programmatic use:

```{r}
#| eval: false
# Get results without printing
results <- sitrep_shinyelectron(verbose = FALSE)

# Check for issues
if (length(results$system$issues) > 0) {
  cat("System issues found:\n")
  print(results$system$issues)
}

# Check Node.js status
if (results$system$node$installed) {
  cat("Node.js version:", results$system$node$version, "\n")
}
```

## Cache Management

If you encounter persistent issues, clearing the cache may help:

```{r}
#| eval: false
# Clear all cached assets
cache_clear("all")

# Clear only Node.js installations
cache_clear("nodejs")

# Clear npm cache
cache_clear("npm")
```

After clearing the cache, reinstall Node.js:

```{r}
#| eval: false
install_nodejs()
```

## Debugging a Packaged App

The packaged Electron app is silent by default — only warnings and errors
print to the console. To see what is happening during runtime detection,
dependency installation, and server startup, launch the app with the
`SHINYELECTRON_DEBUG` environment variable set to `1`:

```bash
# macOS / Linux
SHINYELECTRON_DEBUG=1 /path/to/YourApp.app/Contents/MacOS/YourApp

# Windows (PowerShell)
$env:SHINYELECTRON_DEBUG = "1"; & "C:\Path\To\YourApp.exe"
```

Messages are prefixed with `[shinyelectron]` and include which runtime was
found, the spawn command for R or Python, port retries, and server ready
events — useful context when filing a bug report.

## Getting Help

If diagnostics show all systems ready but you still encounter issues:

1. **Check the console output** for specific error messages (enable
   `SHINYELECTRON_DEBUG=1` for verbose backend logs)
2. **Verify your Shiny app** runs locally with `shiny::runApp()`
3. **Check file permissions** in your project directory
4. **Report issues** at <https://github.com/coatless-rpkg/shinyelectron/issues>

When reporting issues, include the output of:

```{r}
#| eval: false
sitrep_shinyelectron()
```

## Next Steps

- **[Getting Started](getting-started.html)**: First-time user walkthrough
- **[Configuration](configuration.html)**: Customize your builds
- **[Node.js Management](nodejs-management.html)**: Manage Node.js installations
